---
title: 'Stat 3302: Final Project'
author: "Zhengqi Dong"
date: "3/14/2020"
output:
  html_document: default
  pdf_document: 
    latex_engine: xelatex
---

```{r setup, include=FALSE}
# You can load libraries here if you want
library(tidyverse)
library(broom)
library(readr)
# This allows you to set default behavior for R chunks
knitr::opts_chunk$set(echo = TRUE)
```

Define the function to be used:
```{r}
## Define the logit function.
logit <- function (p) 
{
  log(p / (1 - p))
}

## Define the inverse logit function.
sigmoid <- function (etas) 
{
  exp(etas) / (1 + exp(etas))
}

```

Import data:
```{r include=F, echo=F}
Titanic_train <- read_csv("Titanic.train.csv")
```

Process data:


```{r}
head(Titanic_train)
```


## Scientific question: What sorts of people were more likely to survive the Titanic sinking?

## Plot the graph, and summarize your graphs 

##Step1: Build a simple logistic model (without interaction) 

### a) Produce a table including the pclass factor variable, number of passenger survived(survival=1) in each class, and the total number of passenger in each class(survival=1 or 0)

### b.1) Produce a graphical summary that display the relationship between the probability of survival and the factor variable pclass. 

### b.2) Produce a graphical summary that display the relationship between the logit of the probability of survival and the factor variable pclass. 

### c) Fit a model that estimates the logit of the probability of survival in terms of pclass factor variables. Write out the assumed model and define your notation.

### d) Summarize the fitted model and analysis of deviance table.

### e) Discuss the result and idea, is this model help us to understand the scientific question? (What sorts of people were more likely to survive.)

### f) (Option)Interpret this statistical model in terms of changes of the odds with appropriate confidence interval (including at least one confidence intervals for the quantities that you estimated.)


## Step2: Do the same things in terms of factor variable gender

### g)Compare the factor model from above with the model from Step1, using deviance, AICs, and residual plots, which model do you prefer? Why?

## Step3: Do the same things in terms of factor variable Embarked

## Step4: Do the same things in terms of continuous variable age

## Step5: Do the same things in terms of continuous variable fare

## Conclusion: Compare the models from previous step, using deviance, AICs, and residual plots, which model do you prefer? Why?

## Appendix (The R code you need to support previous argument, not counted in the 5–6 page limit)

