---
title: 'Stat 3302: Final Project'
author: "Zhengqi Dong"
date: "3/14/2020"
output:
  html_document: default
  pdf_document: 
    latex_engine: xelatex
---

## 1. Importing library, defining function, and checking data
```{r setup, include=FALSE}
# Basic package
library(tidyverse)
library(broom)
library(readr)
library(ggplot2)
# Advanced(Fancy) package
library(ggthemes)
library(gridExtra)
# This allows you to set default behavior for R chunks
knitr::opts_chunk$set(echo = TRUE)
```


Import data:
```{r include=F, echo=F}
train <- read.csv("train.csv", stringsAsFactors = F)
test <- read.csv('test.csv', stringsAsFactors = F)
full  <- bind_rows(train, test) # bind training & test data
```

```{r}
str(train)
str(test)
str(full)
```
Our training dataset contains 891 observation and 12 variable; testing dataset contains 418 observation with missed variable “Survived” (That's the variable we want to predict with for testing dataset); full dataset contains 1309 observation and 12 variables.
## 2. EDA

### 2.1 Age vs Survived

```{r}
# Age vs Survived
ggplot(full[1:891,], aes(Age, fill = factor(Survived))) + 
  theme_few()+
  geom_histogram(bins=30) + 
  ggtitle("Age vs Survived") + 
  scale_fill_discrete(name = "Survived") # For the label in the right
```
Note: 1)People who are in range between 20 to 50 are less likely being survived, and the people who below age 10 are very likely being survived. 2)The distribution of age looks like a right skewed.

### 2.2 Sex Vs Survive
```{r}
# Sex vs Survived
ggplot(full[1:891,], aes(Sex, fill = factor(Survived))) + 
  geom_bar(stat = "count", position = 'dodge')+
  xlab("Sex") +
  ggtitle("Sex vs Survived") +
  geom_label(stat='count',aes(label=..count..))+
  scale_fill_discrete(name = "Survived") # For the label in the right
```
Note: 1)female is more likely being survived than male, femail survived rate roughly 75%, and male is roughly 16.7%, so almost 5 time greater!

### 2.3 Age Vs Sex Vs Survived
```{r}
#Sex vs Survived vs Age 
ggplot(full[1:891,], aes(Age, fill = factor(Survived))) + 
  geom_histogram(bins=30) + # bins: controls the width of bar, so larger the thinner. You can use geom_bar() if you don't want to specify it!
  xlab("Age") +
  facet_grid(.~Sex)+
  ggtitle("Age vs Sex vs Survived") +
  # geom_label(stat='count',aes(label=..count..)) +
  scale_fill_discrete(name = "Survived") # For the label in the right

```
Note: 1) Again, female is more likely being survived than male. 2) The differences between the number of peoples survived and not survived is largest at roughly age 20to30, and this is true for both female and male.

### 2.4. Pclass vs Sex
```{r}
# geom_bar vs geom_hist: 
#   - Bar charts provide a visual presentation of categorical data
#   - Histograms are used to plot the distribution of data

# Pclass vs Sex Vs Survived
p1 <- ggplot(full[1:891,], aes(Pclass, fill = factor(Survived))) + 
  geom_bar(stat='count') + 
  xlab("Pclass") +
  facet_grid(.~Sex)+
  ggtitle("Pclass vs Sex vs Survived") +
  geom_label(stat='count',aes(label=..count..)) +
  scale_fill_discrete(name = "Survived") # For the label in the right

p2 <- ggplot(full[1:891,], aes(x = Pclass, fill = factor(Survived))) +
  geom_bar(stat='count', position='fill') +
  labs(x = 'Passenger class', y= "Percent") + 
  facet_grid(.~Sex) +
  theme(legend.position="none")
grid.arrange(p1, p2, ncol=2)
```
Note: 1) female is more likely of being survived than male in average. 2)In female group, majority passengers in class 1 and class 2 are survived, and more people in classed died. However, in male group, the survived rate in class 2 (~18.7%) just as bad as class 3(~15.67%). 


## 3. Processing data and Further EDA
Notices, there are some useful infromation in assenger name, what is it? 1)the passenger title!(e.g. Ms, Miss, Mrs..) So, with that being said, we can use to ask some question like, is there any relationship between the passenger title and probability of survived?
2) The surname can be useful as well. It allow us to use "surname" to represent a families. Using the sum of "sibsp" and "parch" as a new variable, we can evaluate the relationship between family size and the probabilty of being survived!

### 3.1 Feature Engineer work:
```{r}
# Grab title from passenger names
full$Title <- gsub('(.*, )|(\\..*)', '', full$Name)

cat("Show title counts by sex:")
table(full$Sex, full$Title)
# kable(table(full$Sex, full$Title)) # A fance table for html presentation

# Titles with very low cell counts to be combined to "rare" level
rare_title <- c('Dona', 'Lady', 'the Countess','Capt', 'Col', 'Don', 
                'Dr', 'Major', 'Rev', 'Sir', 'Jonkheer')

# Also reassign mlle, ms, and mme accordingly
full$Title[full$Title == 'Mlle']        <- 'Miss' 
full$Title[full$Title == 'Ms']          <- 'Miss'
full$Title[full$Title == 'Mme']         <- 'Mrs' 
full$Title[full$Title %in% rare_title]  <- 'Rare Title'

cat("\nShow title counts by sex after merged title has very few count in the data: ") 
table(full$Sex, full$Title)

# Finally, grab surname from passenger name
full$Surname <- sapply(full$Name,  
                      function(x) strsplit(x, split = '[,.]')[[1]][1])
cat("\nHead of full$Surname:\n")
head(full$Surname)
```

### 3.2 Title Vs Survived
```{r}
ggplot(full[1:891,], aes(x = Title, fill = factor(Survived))) +
  geom_bar(stat='count', position='stack') +
  ggtitle("Title Vs Survived") +
  geom_label(stat='count',aes(label=..count..)) +
  scale_fill_discrete(name = "Survived") # For the label in the right
```
Note: we see the Mr. "Mr" are died pretty badly, which proved our previous observation that male are less likely being survived than female.

### 3.3 Family size Vs Survived
Create variable Fsize, which is sum of the number of siblings/spouses and number of chldren/parens and one(The person himself)
```{r}
# Create a family size variable including the passenger themselves
full$Fsize <- full$SibSp + full$Parch + 1

# Create a family variable 
full$Family <- paste(full$Surname, full$Fsize, sep='_')
head(full$Family)
```


```{r}
# Use ggplot2 to visualize the relationship between family size & survival
ggplot(full[1:891,], aes(x = Fsize, fill = factor(Survived))) +
  geom_bar(stat='count', position='dodge') +
  scale_x_continuous(breaks=c(1:11)) +
  ggtitle("Family size Vs Survived") + 
  labs(x = 'Family Size') +
  # geom_label(stat='count',aes(label=..count..)) +
  scale_fill_discrete(name = "Survived") # For the label in the right

```
Note: By comparing the "family size" and "Survived", we noticed the singleton, familes sizes 1, and large families (size > 5) are less likely being survived than the family with size between 2 and 4. Keep this in mind, that might be something we want to use in building our regression model.







## 4. SLLR pairwised: 
Produce a table including the pclass factor variable, number of passenger survived(survival=1) in each class, and the total number of passenger in each class(survival=1 or 0)

Response variable:
  - survived (0 == died, 1 == survived)
Explanatory variables/covaraite of interest:
 - Pclass
 - Sex
 - Age
 - Fsize
 
Define the function to be used:
```{r}
## Define the logit function.
logit <- function (p) 
{
  log(p / (1 - p))
}

## Define the inverse logit function.
sigmoid <- function (etas) 
{
  exp(etas) / (1 + exp(etas))
}

```

### 4.1 Redefined the Fsize to factor: 
Because the purpose of this project is to build a logistic regression model(binary), in order to make the varaible "Fsize" to be more useful and easier to deal with, we need to convert it to factor! (as well as other categorical variables. Factor just a nice data type in R, that is design for categorical variable.)


```{r}
train <- full[1:891,]
## define 'Survived' to be 1 if any passenger survived; 0 if died
train$Survived <- as.numeric(train$Survived == 1)


## define the variable 'Sex'
## is 0 if Sex is light medium or medium.
## is 1 if color is dark medium or dark.
# Sex <- factor(ifelse(crabs$color <= 2, "not dark", "dark"))
train$Sex <- factor(train$Sex)

## Redefine the Pclass factor variable ordered as
## 1: Class_1, 2: Class_2, 3: Class_3, Otherwise: Error.
## (factors by default are ordered alphabetically)
train$Pclass <-
  factor(ifelse(train$Pclass==1, "Class_1",
                ifelse(train$Pclass==2, "Class_2",
                       ifelse(train$Pclass==3, "Class_3", "Error"))),
         levels=c("Class_1", "Class_2", "Class_3"))

## Redefine the Fsize factor variable ordered as
## 1: solo, 2: double, 3: Fsize_3, 4: Fsize_4, 5: large_family
## (factors by default are ordered alphabetically)
# train$Fsize <-
#   factor(ifelse(train$Fsize==1, "solo",
#             ifelse(train$Fsize==2, "double",
#               ifelse(train$Fsize==3, "Fsize_3",
#                 ifelse(train$Fsize==4, "Fsize_4", "large_family")))),
#          levels=c("solo", "double", "Fsize_3", "Fsize_4", "large_family"))

# Alternatively: 
train$Fsize[train$Fsize>=5] <- 'large_family' # THis must go first, otherwise it won't work
train$Fsize[train$Fsize==1] <- 'solo'
train$Fsize[train$Fsize==2] <- 'double'
train$Fsize[train$Fsize==3] <- 'Fsize_3'
train$Fsize[train$Fsize==4] <- 'Fsize_4'
train$Fsize <- as.factor(train$Fsize)
```


```{r}
# Fsize Vs Survived
ggplot(train[!is.na(full$Survived),], aes(x = Fsize, fill = factor(Survived))) +
  geom_bar(stat='count', position='dodge') +
  ggtitle("Family size Vs Survived") + 
  labs(x = 'Family Size') +
  # geom_label(stat='count',aes(label=..count..)) +
  scale_x_discrete (limits = c('solo', 'double', 'Fsize_3', 'Fsize_4', 'large_family')) +
  scale_fill_discrete(name = "Survived") # For the label in the right
```

### 4.2 SLLR on Survived~Age
```{r eval=FALSE, include=FALSE}
rounded_age <- round(train$Age*2)/2
count_table <- table(rounded_age, train$Survived)
prob_survived <- round(count_table / apply(count_table, 1, sum), 3)[,2]
```

```{r eval=FALSE, include=FALSE}

plot(sort(unique(rounded_age)), logit(prob_survived),
     xlab="weight (to nearest 0.5kg)", ylab="proportion")
```


```{r}
Survived_model1 <- glm(train$Survived ~ train$Age, family=binomial)
summary(Survived_model1)
anova(Survived_model1, test="Chisq")
```
p-value(>|Z|) tells us, the coefficient for age is sig different from zero, and has a negative relationship with survived. The pr(>Chi) tells us it's useful to include Age into our model!

### 4.3 SLLR on Survived~Sex

```{r}
# show the table of counts
count_table <- table(train$Sex, train$Survived)
count_table
```

```{r}
# show the table of proportions: p_ij = r_ij/(r_i + r_j): 
round(count_table / apply(count_table, 1, sum), 3)
```
En, looks like female has a high probability of being survived!

```{r}
# fit the glm with Sex: 
Survived_model1 <- glm(train$Survived ~ train$Sex, family=binomial)
summary(Survived_model1)
anova(Survived_model1, test="Chisq")
```
p-value(>|Z|) tells us, the coefficient for age is sig different from zero, and the expected probability of being survived for male is $e^{-2.5137} * 100% = 8.097% $ less than the female in average. The pr(>Chi) tells us it's useful to include sex into our model!

### 4.4 SLLR on Survived~Pclass
```{r}
# show the table of counts
count_table <- table(train$Pclass, train$Survived)
count_table
```

```{r}
# show the table of proportions: p_ij = r_ij/(r_i + r_j): 
round(count_table / apply(count_table, 1, sum), 3)
```

```{r}
# fit the glm with Sex: 
Survived_model1 <- glm(train$Survived ~ train$Pclass, family=binomial)
summary(Survived_model1)
anova(Survived_model1, test="Chisq")
```


### 4.5 SLLR on Survived~Fsize
```{r}
# show the table of counts
count_table <- table(train$Fsize, train$Survived)
count_table
```

```{r}
# show the table of proportions: p_ij = r_ij/(r_i + r_j): 
round(count_table / apply(count_table, 1, sum), 3)
```

```{r}
# fit the glm with Sex: 
Survived_model1 <- glm(train$Survived ~ train$Fsize, family=binomial)
summary(Survived_model1)
anova(Survived_model1, test="Chisq")
```

### 4.6 SLLR on Survived~Age + Sex + Pclass + Fsize
```{r}
Survived_model1 <- glm(train$Survived ~ train$Age + train$Sex + train$Pclass + train$Fsize, family=binomial)
summary(Survived_model1)
anova(Survived_model1, test="Chisq")
```


## 5. SLLP with interaction terms
```{r}
Survived_model1 <- glm(train$Survived ~ train$Age * train$Sex * train$Pclass * train$Fsize, family=binomial)
summary(Survived_model1)
anova(Survived_model1, test="Chisq")
```


## 6. Making prediction


## Scientific question: What sorts of people were more likely to survive the Titanic sinking?

## Step 1: EDA

### b.1) Produce a graphical summary that display the relationship between the probability of survival and the factor variable pclass. 

### b.2) Produce a graphical summary that display the relationship between the logit of the probability of survival and the factor variable pclass. 

### c) Fit a model that estimates the logit of the probability of survival in terms of pclass factor variables. Write out the assumed model and define your notation.

### d) Summarize the fitted model and analysis of deviance table.

### e) Discuss the result and idea, is this model help us to understand the scientific question? (What sorts of people were more likely to survive.)

### f) (Option)Interpret this statistical model in terms of changes of the odds with appropriate confidence interval (including at least one confidence intervals for the quantities that you estimated.)


## Step2: Do the same things in terms of factor variable gender

### g)Compare the factor model from above with the model from Step1, using deviance, AICs, and residual plots, which model do you prefer? Why?

## Step3: Do the same things in terms of factor variable Embarked

## Step4: Do the same things in terms of continuous variable age

## Step5: Do the same things in terms of continuous variable fare

## Conclusion: Compare the models from previous step, using deviance, AICs, and residual plots, which model do you prefer? Why?

## Appendix (The R code you need to support previous argument, not counted in the 5–6 page limit)

